On the specific recursive structure of iterated parametricity

### 2025-11-11 paper/paper.tex
Postit
Experimenting abbreviating depscohs and overline{deps} already in the mutual definition of restrframe and cohftype + fixes

@@

Thanks to $\restrptype[n][0,...,p-1]$, we can mutually define $\restrf[n+1][0,...,p]$ and $\cohftype[n][0,...,p-1]$ from ``full dependencies'', that is from $\framep[n][0,...,n]$, $\painting[n][0,...,n]$, $\restrf[n][0,...,n]$ and some $E:\framep[n+1][n](\deps[n][0,...,n-1]) \rightarrow \U$, using abbreviations in passing for $\restrl$ and $\restrltype$:

@@

                \restrp[n][0,...,p-1]:\restrptype[n][0,...,p-1]
                (\fulldeps[n])    \\
                \cohf[n][0,...,p-1]:\cohftype[n][0,...,p-1]

@@

              \end{array}\right) :

@@

                \framep[n+1][0,...,p](\fulldeps.\deps[n][0,...,p])   \\
                \painting[n+1][0,...,p](\fulldeps[n]) \\

@@

            \end{array}\right)                                                       & \defeq & (\unitpoint,\lambda q. \lambda\epsilon. \lambda \_.\unitpoint) \\

@@

@@

                  \cohf[n][0,...,p-1]   \\

@@

                    \cohf[n][0,...,p-1]   \\
                  \end{array}\right)(d), \restrl[n][p][q][\epsilon]
                  \end{array}\right)(d)(l)\right)     \\

@@

@@

                \framep[n+1][p](\fulldeps.\deps[n][0,...,p-1])          \\
                \painting[n+1][0,...,p](\fulldeps[n]) \\
                \restrf[n+1][0,...,p]

@@

                  \cohf[n][0,...,p-1] \\

@@

                \cohf[n][0,...,p-1] \\

@@

              \restrp[n][0,...,p]:\restrptype[n][0,...,p](\fulldeps[n]) \\
              \cohf[n][0,...,p]:\cohftype[n][0,...,p]

@@

@@

    \Pi q \leq n. \Pi \epsilon. \Pi d: \framep[n+1][p].\\
    \left(
            \framep[n][p](\fulldeps.\deps[n][0,...,p]) \\
            \painting[n][p](\fulldeps[n])              \\
            \restrf[n][p]
          \end{array}\right)(d) \rightarrow \\

@@

                  \cohf[n][0,...,p-1]
        \end{array}\right)(d))

@@

              \restrp[n][0,...,p]:\restrptype[n][0,...,p](\fulldeps[n]) \\
              \cohf[n][0,...,p]:\cohftype[n][0,...,p]
    \restrltype[n][p] \left(\begin{array}{l}
    \right)                                                                                                                                                                           \\

@@

    \lambda q\leq n. \lambda \epsilon.\,\lambda d.\lambda l.\overrightarrow{\cohf[n][p][q][q][\epsilon][\omega](d)}{\restrp[n][p][q][\epsilon](\restrf[n][p][q][\epsilon]\left(\begin{array}{l}
                  \cohf[n][0,...,p-1]
                \end{array}\right)(d))(l \omega)}                          \\

@@

At this new stage, it is convenient to use a new abbreviation to combine sequences of $\deps[n][0,..,n]$, $\restrp[n][0,...,p-1]$ and $\cohf[n][0,...,p-1]$ in one block, which we call $\depscohstype[n][0,...p-1]$:

\begin{align*}
  \depscohstype[n][0,..,p-1] \defeq
  \left(\begin{array}{lcl}
            \fulldeps[n][0,...,n] & : & \fulldepstype[n][0,...,n] \\
            \restrp[n][0,...p-1]  & : & \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]   & : & \cohftype[n][0,...,p-1]   \\
          \end{array}\right)
\end{align*}
so that the type of $\restrf[n+1][0,...,p]$ can be rephrased as $\Pi \depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1]. \restrftype[n+1][0,...,p]
      \left(\begin{array}{l}
                \framep[n+1][0,...,p](\fulldeps.\deps[n][0,...,p])   \\
                \painting[n+1][0,...,p](\fulldeps[n]) \\
              \end{array}\right)$, and similarly for $\restrltype$ and $\restrl$. Later, we shall write $\depscohs.\deps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of an element $\depscohs$ in $\depscohstype$. By abuse of notation, we also write $E$ for an inhabitant of $\depscohstype[n][0,..,0-1]$ (standing for a non trivial inhabitant of $\fulldeps[n]$).

It is also convenient to build $\deps$ from $\depscohs$ at the
previous level:
\begin{align*}
  \overline{\deps}(\depscohs[n][0,...,p-1]:\depscohstype[n][0,...,p-1]) & :      & \depstype[n+1][0,...,p] \\
  \overline{\deps}(\depscohs)               & \defeq &
  \left(\begin{array}{l}
            \framep[n+1][0,...,p](\depscohs.\fulldeps.\deps[n][0,...,p-1])   \\
            \painting[n+1][0,...,p](\depscohs.\fulldeps[n]) \\
            \restrf[n+1][0,...,p](\depscohs[n][0,...,p-1])         \\
          \end{array}\right)             \\
\end{align*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][0,...,p-1] & (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \U                                      \\
    \cohptype[n][0,...,0-1]        & \unitpoint                                        & \defeq & \unittype                               \\
    \cohptype[n][0,...,(p+1)-1] & (\fulldepscohs[n])                           & \defeq & \cohptype[n][0,...,p-1](\fulldepscohs[n]) \times \left(
    \begin{array}{l}
        \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi d:\framep[n+2][p](\overline{\deps}(\fulldepscohs.\depscohs[n][0,...,p-1])).\,\Pi c:\painting[n+2][p](\overline{\fulldeps}(\fulldepscohs[n]))\\
        \fulldepscohs.\depscohs.\restrp[n][p][q][\epsilon](\restrf[n+1][p](d))(\restrp[n+1][p][r][\omega]
        (\fulldepscohs[n])(d)(c)) =                         \\
        \fulldepscohs.\depscohs.\restrp[n][p][r][\omega](\restrf[n+1][p](d))(\restrp[n+1][p][q+1][\epsilon]
        (\fulldepscohs[n])(d)(c))
      \end{array}
    \right)                                                                                                            \\
  \end{array}
\end{equation*}

@@

                  \restrp[n+1][0,...,p-1] \\

@@

              \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi \theta  \\

>>>>

TODO: talk about unfoldPaintingProj.

@@

Thanks to $\restrptype[n][0,...,p-1]$, we can mutually define $\restrf[n+1][0,...,p]$ and $\cohftype[n][0,...,p-1]$ from $\framep[n][0,...,n]$, $\painting[n][0,...,n]$, $\restrf[n][0,...,n]$, $\restrptype[n][0,...,p-1]$ and some $E:\framep[n+1][n](\deps[n][0,...,n-1]) \rightarrow \U$, abbreviating these dependencies and the dependency in a $\cohf[n][0,...,p-1]$ as $\depscohttype[n][0,...,p-1]$, and using also abbreviations in passing for $\restrl$ and $\restrltype$. It is also convenient to build $\deps[n+1][0,...,p]$ from $\depscohs[n][0,...,p-1]$, so that we can now build $\framep[n+2][p]$ one step further. Thus, we also include this abbreviation in the definition:

@@

    \depscohstype[n][0,..,p-1] & \defeq &
    \left(\begin{array}{lcl}
              \fulldeps[n][0,...,n] & : & \fulldepstype[n][0,...,n] \\
              \restrp[n][0,...p-1]  & : & \restrptype[n][0,...,p-1](\fulldeps[n]) \\
              \cohf[n][0,...,p-1]   & : & \cohftype[n][0,...,p-1]

@@

    \overline{\deps}(\depscohs[n][0,...,p-1]:\depscohstype[n][0,...,p-1]) & :      & \depstype[n+1][0,...,p] \\
    \overline{\deps}(\depscohs)               & \defeq &
              \framep[n+1][0,...,p](\depscohs.\fulldeps.\deps[n][0,...,p-1])   \\
              \painting[n+1][0,...,p](\depscohs.\fulldeps[n]) \\
              \restrf[n+1][0,...,p](\depscohs[n][0,...,p-1])         \\
      (\depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1]) :

@@

                \framep[n+1][0,...,p](\depscohs.\fulldeps.\deps[n][0,...,p])   \\
                \painting[n+1][0,...,p](\depscohs.\fulldeps[n]) \\

@@

    (E)
    & \defeq & (\unitpoint,\lambda q. \lambda\epsilon. \lambda \_.\unitpoint) \\

@@

    (\depscohs[n][0,..,p])                                                       & \defeq &

@@

              (\depscohs[n][0,..,p-1]),             \\

@@

              (\depscohs[n][0,..,p-1])(d), \restrl[n][p][q][\epsilon]
              (\depscohs[n][0,..,p])(d)(l)\right)     \\

@@

    }                                                                                                                  n                                                               \\

@@

      (\overline{\deps}\left(\begin{array}{l}

@@

                    \end{array}\right)          .                         \\

@@

@@

      (\depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1])
      & :      & \U                                                                               \\

@@

      (\depscohs[n][0,..,p-1])

@@

    \Pi q \leq n. \Pi \epsilon. \Pi d: \framep[n+2][p](\overline{\deps}(\depscohs[n][0,..,p-1])).\, \\
        \layer[n+1][p](\overline{\deps}(\depscohs[n][0,..,p-1])(d) \rightarrow \\

@@

        (\depscohs[n][0,..,p-1])(d))

@@

      (\depscohs[n][0,..,p]:\depscohstype[n][0,..,p]) & :      &
    \restrltype[n][p] (\depscohs[n][0,..,p])                                                   \\

@@

    (\depscohs[n][0,..,p])                                                     & \defeq &
    \lambda q\leq n. \lambda \epsilon.\,\lambda d.\lambda l.\overrightarrow{\cohf[n][p][q][q][\epsilon][\omega](d)}{\restrp[n][p][q][\epsilon](\restrf[n][p][q][\epsilon](\depscohs[n][0,..,p-1])(d))(l \omega)}                          \\

@@

By abuse of notation, we also write $E$ for an inhabitant of $\depscohstype[n][0,..,0-1]$ (standing for a non trivial inhabitant of $\fulldeps[n]$).

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][0,...,p-1] & (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \U                                      \\
    \cohptype[n][0,...,0-1]        & \unitpoint                                        & \defeq & \unittype                               \\
    \cohptype[n][0,...,(p+1)-1] & (\fulldepscohs[n])                           & \defeq & \cohptype[n][0,...,p-1](\fulldepscohs[n]) \times \\
    & & & \;\;\;\;\left\{
    \begin{array}{l}
        \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi d:\framep[n+2][p](\overline{\deps}(\fulldepscohs.\depscohs[n][0,...,p-1])).\,\Pi c:\painting[n+2][p](\overline{\fulldeps}(\fulldepscohs[n]))\\
        \fulldepscohs.\depscohs.\restrp[n][p][q][\epsilon](\restrf[n+1][p](d))(\restrp[n+1][p][r][\omega]
        (\fulldepscohs[n])(d)(c)) =                         \\
        \fulldepscohs.\depscohs.\restrp[n][p][r][\omega](\restrf[n+1][p](d))(\restrp[n+1][p][q+1][\epsilon]
        (\fulldepscohs[n])(d)(c))
      \end{array}
    \right.                                                                                                           \\
  \end{array}
\end{equation*}

@@

@@

              \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi \theta.\,  \\
